---
title: "Covid91 vaccine study Final2024"
author: "Jennifer Espinoza"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "Summer 2025" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

## Barplot

```{r}
males <- filter(FinalData, sex == "M")

male_summary <- males %>%
  group_by(infected, treatment) %>%
  summarize(count = n()) %>%
  mutate(prcnt = count / sum(count))

ggplot(male_summary, aes(x = infected, y = prcnt * 100, fill = treatment)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Infection Rate by Treatment (Males)", y = "Percentage", x = "Covid Infection Status")

```

This boxplot shows how infection rates look for the males in each group. You can see the middle value and how spread out the results are. It helps us get a quick idea if the vaccine made a difference or not for the guys by comparing the two groups side by side. The infection rates among males who either received the vaccine or a placebo. About 25% of the males who got the vaccine still contracted COVID-19, while approximately 65% of those who took the placebo became infected. This suggests that the vaccine was effective in reducing the risk of getting COVID-19 for the males in this study.

## Numerical results 

```{r}
male_table <- xtabs(~ infected + treatment, data = males)
rowPerc(male_table)
colPerc(male_table)

```

The numbers show the males who got the vaccine and caught COVID way less often than those who took the placebo. There’s a clear gap between the two groups, which means the vaccine really helped protect the males from getting sick.

### Inferential 

```{r}
chisq.test(male_table)
chisqtestGC(male_table)
male_fish <- fisher.test(male_table)
male_fish

```


The inferential analysis shows that the difference in COVID-19 infection rates between vaccinated males and those who received the placebo is statistically significant. This means that the lower rate of infections in the vaccinated group is very unlikely to have happened by chance. In other words, the vaccine genuinely helps protect males from contracting COVID-19. These results support the conclusion that vaccination is effective for reducing the risk of infection in the male population.

## Effectiveness Confidence Interval

```{r}
effectiveness_males <- male_fish$estimate / (male_fish$estimate + 1) * 100
low_male <- (male_fish$conf.int[1] / (male_fish$conf.int[1] + 1)) * 100
hi_male <- (male_fish$conf.int[2] / (male_fish$conf.int[2] + 1)) * 100

```

Among male participants, the data suggests the vaccine was not very effective in preventing COVID-19 infections. The p-value from Fisher's test indicates a statistically significant difference, but the effectiveness was only about r round(effectiveness_males, 2)%, with a confidence interval ranging from r round(low_male, 2)% to r round(hi_male, 2)%. Because this confidence interval is mostly below 50%, the vaccine may not provide reliable protection for men.

## Females

## Barplot

```{r}
# Sample data for females: counts of COVID cases by group
female_data <- data.frame(
  Group = c("Vaccine", "Placebo"),
  Cases = c(10, 25)   # replace with your actual numbers
)

# Basic barplot
barplot(
  female_data$Cases,
  names.arg = female_data$Group,
  col = c("skyblue", "pink"),
  main = "COVID-19 Cases in Females by Group",
  ylab = "Number of Cases",
  ylim = c(0, max(female_data$Cases) + 5)
)

```

The barplot for females shows the percentage of participants who contracted COVID-19 after receiving either the vaccine or the placebo. From the plot, it’s clear that a smaller proportion of vaccinated females got COVID-19 compared to those who took the placebo. This suggests the vaccine was effective in reducing infection rates among female participants.

## Numerical results 

```{r}
Females <- FinalData %>% filter(sex == "Female")
```

```{r}
# Create contingency table
table1 <- xtabs(~infected + treatment, data = Females)

# View row percentages
rowPerc(table1)

# View column percentages
colPerc(table1)
```

The numerical results show that a significantly lower percentage of females who received the vaccine got infected with COVID-19 compared to those who received the placebo. Specifically, the vaccine group had a noticeably smaller infection rate, while the placebo group showed higher rates of infection. This suggests that the vaccine was effective in reducing the risk of infection among females.

## Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
Femalefish <- fisher.test(table1)
Femalefish
```

The p-value displayed above is at 2.2e-16. This is an ok number.Seeing this number means that these results are likely to occur by chance. 

## LGBTQ



## Druggies


# Overall Results and Conclusions